Fast Hyper-Parameter Selection for Graph Kernels via Subsampling and Multiple Kernel Learning

### 2016-11-25 paper.tex
experiments

% My commands
\newcommand{\Real}{\mathbb{R}}
\newcommand{\Natu}{\mathbb{N}}
\newcommand{\dd}{{\bf d}}
\newcommand{\xx}{{\bf x}}
\newcommand{\yy}{{\bf y}}
\newcommand{\zz}{{\bf z}}
\newcommand{\ff}{{\bf f}}
\newcommand{\ww}{{\bf w}}
\newcommand{\vv}{{\bf v}}
\newcommand{\AAA}{{\bf A}}
\newcommand{\CC}{{\bf C}}
\newcommand{\DD}{{\bf D}}
\newcommand{\EE}{{\bf E}}
\newcommand{\GG}{{\bf G}}
\newcommand{\KK}{{\bf K}}
\newcommand{\YY}{{\bf Y}}
\newcommand{\MM}{{\bf M}}
\newcommand{\HH}{{\bf H}}
\newcommand{\II}{{\bf I}}
\newcommand{\XX}{{\bf X}}
%\newcommand{\NN}{{\bf N}}
\newcommand{\cU}{{\cal U}}
\newcommand{\cI}{{\cal I}}
\newcommand{\cR}{{\cal R}}
\newcommand{\cN}{{\cal N}}
\newcommand{\cQ}{{\mathcal Q}}
\newcommand{\1}{{\bf 1}}
\newcommand{\Exp}{\mathbb{E}}
\newcommand{\labelfun}{L}
\newcommand{\node}{v}
\newcommand{\cK}{{\cal K}}
%\newtheorem{lemma}{Lemma}
%\newtheorem{theorem}{Theorem}
%\newtheorem{corollary}{Corollary}
%\newtheorem{proposition}{Proposition}
%\newtheorem{definition}{Definition}
%\newtheorem{remark}{Remark}
\newcommand{\cC}{{\cal C}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\ggamma}{\pmb{\gamma}}
\newcommand{\aalpha}{\pmb{\alpha}}
\newcommand{\bbeta}{\pmb{\beta}}
\newcommand{\eeta}{\pmb{\eta}}
\newcommand{\ssigma}{\pmb{\sigma}}
\newcommand{\llambda}{\pmb{\lambda}}
\newcommand{\mmu}{\pmb{\mu}}
\newcommand{\pphi}{\pmb{\phi}}
\newcommand{\res}[2]{$#1_{\pm#2}$}

@@

In this paper we focus on graphs, i.e. we consider each example to be a different graph.
A graph is a tuple $G=(V_G,E_G,L_G)$, where $V_G=\{v_1,\ldots,v_n\}$ is the set of vertices, $E_G=\{(u,v) | u,v \in V_G\}$ is the set of edges (with $|E_G|=m$), and $L_G: V_G \rightarrow \Sigma$ is a function mapping each vertex to a discrete label in a fixed alphabet $\Sigma$.
%A graph is undirected if $(u,v)\in E_g \Rightarrow (v,u) \in E_G$, otherwise it is directed.
In this setting, there are several graph kernels defined in literature based on random walks~\cite{Kashima03marginalizedkernels,Mahe2004} ($O(n^3)$ computational complexity), shortest paths~\cite{Kriegel05shortestpath} ($O(n^4)$ computational complexity), subgraphs up to a fixed size $h$~\cite{Shervashidze2009} ($O(n^h)$ complexity). The problem with these kernels is the high computational complexity, that makes them inapplicable to several real-world datasets.
Recently, different almost linear time graph kernels have been defined in literature \cite{Heinonen2009,NIPS2009_0533,Costa2010,Shervashidze2011,Dasan2012,DaSanMartino2016}, and these are the ones we focus on in this paper.
In the following we will give some details about the considered alternatives.
Each kernel restricts to specific kinds of subgraphs, up to a certain dimension (that, in general, is a kernel parameter).
Note that it is out of the scope of this paper to give an extensive comparison among the various graph kernels.

The Weisfeiler-Lehman Fast Subtree kernel (WL)~\cite{NIPS2009_0533}  counts the number of identical subtree patterns obtained by subtree-walks up to height $h$ (an hyper-parameter) in the two input graphs. The complexity of the kernel is $O(|E|h)$. 
%While being fast to compute, the kernel may lack of expressiveness for some tasks given that the number of non-zero features generated by one graph is at most $|V|h$. 
Note that the subtree-walks extracted by the kernel differ from  subtree structures because a node can appear multiple times in the same subtree-walk.

@@

These algorithms are supported by several theoretical results that bound the \emph{estimation error} (i.e. the difference between the true error and the empirical margin error). These bounds exploit the \emph{Rademacher complexity} applied to the combination of kernels \cite{Maurer2012,Srebro2006,Cortes2009c,Hussain2011,Hussain2011a}. %,Kakade2009,Micchelli2005,Kloft2012

Existing MKL approaches can be divided in two main categories. In the first category, \emph{Fixed or Heuristic}, some fixed rule is applied to obtain the combination. They usually get results scalable with respect to the number of kernels combined but their effectiveness critically depends on the domain at hand. They use a parameterized combination function and find the parameters of this function generally by looking at some measure obtained from each kernel separately,  often giving a suboptimal solution (since no information sharing among the kernels is exploited).

On the other hand, the \emph{Optimization based} approaches learn the combination parameters by solving a single optimization problem directly integrated in the learning machine (e.g. structural risk based target function) or formulated as a different model (e.g. alignment, or other kernel similarity maximization) \cite{Rakotomamonjy2008,Bach2004,Varma2009}. %,Kloft2011,SorenSonnenburg2006,Vishwanathan2010,Xu2010

@@

The objective function in Eq. \ref{eq:easymklotig} can be interpreted as the dual problem of a regularized empirical objective function using the kernel $\sum_{r=1}^R \eta_r \KK_r$. This equation is a minimax problem that can be reduced  to a simple quadratic problem with a technical derivation described  in \cite{Aiolli2015}. The solution of the quadratic problem is an optimal $\ggamma^*$ for the original min-max formulation. Due to the particular structure of EasyMKL, it is sufficient to provide the average kernel of all the basic kernels ($\KK^A = \frac{1}{R} \sum_{r=1}^R \frac{\KK_r}{Tr(\KK_r)}$). From $\ggamma^*$, it is easy to obtain the optimal weights for the single basic kernels $\KK_r$ by using the following formula
\begin{equation}
\label{eq:easyeta}
	\eta_r = \gamma^{*T} \YY \,\, (\KK_r/Tr(\KK_r)) \,\, \YY \gamma^*, \,\,\,\, \forall r=1,\dots,R.
\end{equation}

@@

Firstly, we performed some experiments to observe how many samples are required to make $\eta$ and $\eta_s$ similar.
The metric used for comparison is the jaccard similarity score. Indeed, in our context two weights vectors are similar is they provide the same top kernels.
These experiments are summarized in \figurename\ \ref{fig:weights} and \ref{fig:jaccard}.

% \begin{figure}[!htb]
% \centering
% \includegraphics[width=\textwidth]{img/weights_GDD.png}
% \caption{The weights vector learned by EasyMKL using $50\%$ and $25\%$ of training data. Blue regions represents the comparison with the weights vector calculated using all data.}
% \label{fig:weights}
% \end{figure}

\begin{figure}[!htb]
\centering
\subfloat[a]{\includegraphics[width=0.42\textwidth]{img/CPDB_jaccard.png}}
\subfloat[b]{\includegraphics[width=0.42\textwidth]{img/GDD_jaccard.png}}
\caption{Average jaccard values and standard deviation in 50 runs between the indices of top kernels using a subset of training data and using the whole set.}
\label{fig:jaccard}
\end{figure}

In the first step of our method, we learn the weights vector $\eta_{10}$ using a subset $(10\%)$ of training data and the whole set of kernels. To do this, we perform a 10-fold cross validation using EasyMKL to combine kernels and SVM as base learner. The hyperparameters are $(\lambda,C)$. Then we fit the model with the best combination and chose the top $10\%$ kernels.\\
In the latter step we calculate the top kernel matrices and combine it to get the kernel. We propose three methods to perform this combination:
\begin{itemize}
\item{}use the EasyMKL algorithm;
\item{}calculate the average of kernels;
\item{}use the same weights (rescaled) from the first learning step.
\end{itemize}
For testing, we compared our method with 3 hard baselines, that are:
\begin{itemize}
\item{}validation: we select the kernel which has the best result in validation;
\item{}average: we calculate the kernel as the average of all base kernels;
\item{}combined: we use EasyMKL to combine the whole set of kernels;
\end{itemize}
In all models, we use a SVM as base learner, and we select the hyperparameters using a nested 10-fold CV procedure. The hyperparameters are $\lambda\in\{0,0.1,0.2,\dots,1\}-\{0.5\}$ for models where we use EasyMKL and $C\in\{2^i, i:-3\dots6\}$.
Note that for all baselines we need to calculate all kernel matrices using all of training data.

The datasets used in our experiments are: [???]

@@

\begin{table}[]
\centering
\caption{My caption}
\label{tab:res}
\resizebox{\textwidth}{!}{
\begin{tabular}{l|c|lll|lll}
\hline
dataset & \# train & validation      & all (easy) & all (average) & subset\_easy & subset\_aver & subset\_w \\
\hline
CPDB    & 684 &$0.8753_{\pm0.029}$ & $0.8814_{\pm0.027}$ & $0.8809_{\pm0.029}$ & $0.8768_{\pm0.023}$ & $0.8850_{\pm0.027}$ & $0.8751_{\pm0.023}$ \\
\\
MUTAG   & 188 & \res{0.9023}{0.075} & \res{0.9310}{0.061} & \res{0.9309}{0.060} & \res{0.9292}{0.054} & \res{0.9292}{0.054} & \res{0.9331}{0.053}    \\
\\
CAS     & 4337 & \res{0.9189}{0.012} & - & \res{0.9209}{0.012} & \res{0.9195}{0.11} & \res{0.9161}{0.012} & \res{0.9185}{0.012}    \\
\\
GDD     & 1178 & \res{0.8647}{0.045} & \res{0.8719}{0.039} & \res{0.8675}{0.039} & \res{0.869}{0.041} & \res{0.8713}{0.043} & \res{0.8011}{0.036}    \\
\\
AIDS    & 1503 & \res{0.7797}{0.083} & \res{0.7775}{0.080} & \res{0.7712}{0.083} & \res{0.7845}{0.077} & \res{0.7690}{0.074} & \res{0.7673}{0.073}    \\
\\
NCI1    & 4110 & \res{0.8839}{0.043} & - & \res{0.8851}{0.041} & \res{0.8818}{0.043} & \res{0.8772}{0.047} & \res{0.8811}{0.044}    \\
\\
NCI109 & 4127 & \res{0.8728}{0.025} & - & \res{0.8713}{0.025} & \res{0.8670}{0.024}  & \res{0.8713}{0.025} & \res{0.8680}{0.022}  \\
\\
\hline
\end{tabular}}
\end{table}

>>>>

% My commands
\newcommand{\Real}{\mathbb{R}}
\newcommand{\Natu}{\mathbb{N}}
\newcommand{\dd}{{\bf d}}
\newcommand{\xx}{{\bf x}}
\newcommand{\yy}{{\bf y}}
\newcommand{\zz}{{\bf z}}
\newcommand{\ff}{{\bf f}}
\newcommand{\ww}{{\bf w}}
\newcommand{\vv}{{\bf v}}
\newcommand{\AAA}{{\bf A}}
\newcommand{\CC}{{\bf C}}
\newcommand{\DD}{{\bf D}}
\newcommand{\EE}{{\bf E}}
\newcommand{\GG}{{\bf G}}
\newcommand{\KK}{{\bf K}}
\newcommand{\YY}{{\bf Y}}
\newcommand{\MM}{{\bf M}}
\newcommand{\HH}{{\bf H}}
\newcommand{\II}{{\bf I}}
\newcommand{\XX}{{\bf X}}
%\newcommand{\NN}{{\bf N}}
\newcommand{\cU}{{\cal U}}
\newcommand{\cI}{{\cal I}}
\newcommand{\cR}{{\cal R}}
\newcommand{\cN}{{\cal N}}
\newcommand{\cQ}{{\mathcal Q}}
\newcommand{\1}{{\bf 1}}
\newcommand{\Exp}{\mathbb{E}}
\newcommand{\labelfun}{L}
\newcommand{\node}{v}
\newcommand{\cK}{{\cal K}}
%\newtheorem{lemma}{Lemma}
%\newtheorem{theorem}{Theorem}
%\newtheorem{corollary}{Corollary}
%\newtheorem{proposition}{Proposition}
%\newtheorem{definition}{Definition}
%\newtheorem{remark}{Remark}
\newcommand{\cC}{{\cal C}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\ggamma}{\pmb{\gamma}}
\newcommand{\aalpha}{\pmb{\alpha}}
\newcommand{\bbeta}{\pmb{\beta}}
\newcommand{\eeta}{\pmb{\eta}}
\newcommand{\ssigma}{\pmb{\sigma}}
\newcommand{\llambda}{\pmb{\lambda}}
\newcommand{\mmu}{\pmb{\mu}}
\newcommand{\pphi}{\pmb{\phi}}

\newcommand{\res}[2]{$#1_{\pm#2}$}

\newcommand{\kval}{$K_{\textsc{val}}$}
\newcommand{\ksum}{$K_{\textsc{sum}}$}
\newcommand{\kmkl}{$K_{\textsc{mkl}}$}
\newcommand{\ksmkl}{$K_{\textsc{mkl}}^{sub}$}
\newcommand{\kssum}{$K_{\textsc{sum}}^{sub}$}
\newcommand{\ksw}{$K_{\textsc{w}}^{sub}$}

@@

In this paper we focus on graphs, i.e. we consider each example to be a different graph.
A graph is a tuple $G=(V_G,E_G,L_G)$, where $V_G=\{v_1,\ldots,v_n\}$ is the set of vertices, $E_G=\{(u,v) | u,v \in V_G\}$ is the set of edges (with $|E_G|=m$), and $L_G: V_G \rightarrow \Sigma$ is a function mapping each vertex to a discrete label in a fixed alphabet $\Sigma$.
%A graph is undirected if $(u,v)\in E_g \Rightarrow (v,u) \in E_G$, otherwise it is directed.
In this setting, there are several graph kernels defined in literature based on random walks~\cite{Kashima03marginalizedkernels,Mahe2004} ($O(n^3)$ computational complexity), shortest paths~\cite{Kriegel05shortestpath} ($O(n^4)$ computational complexity), subgraphs up to a fixed size $h$~\cite{Shervashidze2009} ($O(n^h)$ complexity). The problem with these kernels is the high computational complexity, that makes them inapplicable to several real-world datasets.
Recently, different almost linear time graph kernels have been defined in literature \cite{Heinonen2009,Costa2010,Shervashidze2011,Dasan2012,DaSanMartino2016}, and these are the ones we focus on in this paper.
In the following we will give some details about the considered alternatives.
Each kernel restricts to specific kinds of subgraphs, up to a certain dimension (that, in general, is a kernel parameter).
Note that it is out of the scope of this paper to give an extensive comparison among the various graph kernels.

The Weisfeiler-Lehman Fast Subtree kernel (WL)~\cite{Shervashidze2011}  counts the number of identical subtree patterns obtained by subtree-walks up to height $h$ (an hyper-parameter) in the two input graphs. The complexity of the kernel is $O(|E|h)$. 
%While being fast to compute, the kernel may lack of expressiveness for some tasks given that the number of non-zero features generated by one graph is at most $|V|h$. 
Note that the subtree-walks extracted by the kernel differ from  subtree structures because a node can appear multiple times in the same subtree-walk.

@@

These algorithms are supported by several theoretical results that bound the \emph{estimation error} (i.e. the difference between the true error and the empirical margin error). These bounds exploit the \emph{Rademacher complexity} applied to the combination of kernels \cite{Maurer2012,Srebro2006,Cortes2009c}%,Hussain2011,Hussain2011a}. %,Kakade2009,Micchelli2005,Kloft2012

Existing MKL approaches can be divided in two main categories. In the first category, \emph{Fixed or Heuristic}, some fixed rule is applied to obtain the combination. They usually get results scalable with respect to the number of kernels combined but their effectiveness critically depends on the domain at hand. %They use a parameterized combination function and find the parameters of this function generally by looking at some measure obtained from each kernel separately,  often giving a suboptimal solution (since no information sharing among the kernels is exploited).

On the other hand, the \emph{Optimization based} approaches learn the combination parameters by solving a single optimization problem directly integrated in the learning machine %(e.g. structural risk based target function)
or formulated as a different model (e.g. alignment, or other kernel similarity maximization) \cite{Rakotomamonjy2008,Bach2004,Varma2009}. %,Kloft2011,SorenSonnenburg2006,Vishwanathan2010,Xu2010

@@

The objective function in Eq. \ref{eq:easymklotig} can be interpreted as the dual problem of a regularized empirical objective function using the kernel $\sum_{r=1}^R \eta_r \KK_r$. This equation is a minimax problem that can be reduced  to a simple quadratic problem with a technical derivation described  in \cite{Aiolli2015}. %The solution of the quadratic problem is an optimal $\ggamma^*$ for the original min-max formulation. Due to the particular structure of EasyMKL, it is sufficient to provide the average kernel of all the basic kernels ($\KK^A = \frac{1}{R} \sum_{r=1}^R \frac{\KK_r}{Tr(\KK_r)}$). From $\ggamma^*$, it is easy to obtain the optimal weights for the single basic kernels $\KK_r$ by using the following formula
%\begin{equation}
%\label{eq:easyeta}
%	\eta_r = \gamma^{*T} \YY \,\, (\KK_r/Tr(\KK_r)) \,\, \YY \gamma^*, \,\,\,\, \forall r=1,\dots,R.
%\end{equation}

@@

Firstly, we performed some experiments to check how many samples are required to make $\boldsymbol{\eta}$ and $\hat{\boldsymbol{\eta}}$ similar.
In our context two weights vectors are similar if they provide the same top kernels. By the way the metric used for comparison is the jaccard similarity score. \figurename\ \ref{fig:jaccard} show the jaccard score between the indices of top kernels selected using $\boldsymbol{\eta}$ and $\hat{\boldsymbol{\eta}}$ with different threshold and different subsampling dimension.
From other preliminary results, we fix the dimension of subset used in the first step to learn $\hat{\boldsymbol{\eta}}$ as 10\% and the number of top kernels as 10\%.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.42\textwidth]{img/CPDB_jaccard.png}
\includegraphics[width=0.42\textwidth]{img/GDD_jaccard.png}
\caption{Average jaccard values and standard deviation in 50 runs between the indices of top kernels using a subset of training data and using the whole set.}
\label{fig:jaccard}
\end{figure}

To learn $\hat{\boldsymbol{\eta}}$ we chose a random subsampling of training data and perform a 10-fold cross validation using EasyMKL to combine kernels and SVM as base learner. Then we fit the model with the best combination and select the top kernels.
In the second step we calculate the top kernel matrices using the whole training data and combine it to get the kernel. We propose 3 different methods to perform this combination:
(1)\ksmkl: use the EasyMKL algorithm, (2)\kssum: calculate the average of kernels, (3)\ksw: use the same weights (rescaled) from the first learning step.\\
For testing, we compared our method against 3 hard baselines, that are: (a)\kval: the kernel is selected in the nested-cv procedure, (b)\ksum: the kernel is the average of all base kernels, (c)\kmkl: we use EasyMKL to combine the whole set of kernels.\\
In all models, we use a SVM as base learner, and we select the hyperparameters using a nested 10-fold CV procedure. The hyperparameters are $\lambda\in\{0,0.1\dots0.4,0.6\dots1\}$ for models where we use EasyMKL and $C\in\{2^i, i:-3\dots6\}$.
Note that for all baselines we need to calculate all kernel matrices using all of training data.\\
The kernels used in our experiments are: all NSPDK combinations with $h \in \{0,1\dots9\}$ and $d \in \{0,1\dots9\}$, all ODDST combinations with $h \in \{0,1\dots9\}$ and $\lambda \in \{0.5, 0.7,0.8,0.9,1.0,1.1,1.2,1.4,1.6,1.8\}$, all WL with $h \in \{0,1\dots9\}$.

@@

\begin{table}[]
\centering
\caption{My caption}
\label{tab:res}
\resizebox{\textwidth}{!}{
\begin{tabular}{l|c|ccc|ccc}
\hline
dataset & \# train &\kval &\kmkl &\ksum & \ksmkl & \kssum & \ksw \\
\hline
MUTAG   & 188 & \res{0.9023}{0.075} & \res{0.9310}{0.061} & \res{0.9309}{0.060} & \res{0.9292}{0.054} & \res{0.9292}{0.054} & \res{0.9331}{0.053}    \\
&			  & 18m					& 40m				& 8m		& 3m		&4m		&4m\\\hline
CPDB    & 684 &$0.8753_{\pm0.029}$ & $0.8814_{\pm0.027}$ & $0.8809_{\pm0.029}$ & $0.8768_{\pm0.023}$ & $0.8850_{\pm0.027}$ & $0.8751_{\pm0.023}$ \\
&			  & 1h 16m				& 8h 33m		& 20m	& 1h 37m		& 5m		&5m\\\hline
GDD     & 1178 & \res{0.8647}{0.045} & \res{0.8719}{0.039} & \res{0.8675}{0.039} & \res{0.8690}{0.041} & \res{0.8713}{0.043} & \res{0.8011}{0.036}    \\
&			& +24h	& +24h	& +24h	&3h		& 4h 15m		& 4h 15m\\\hline
AIDS    & 1503 & \res{0.7797}{0.083} & \res{0.7775}{0.080} & \res{0.7712}{0.083} & \res{0.7845}{0.077} & \res{0.7690}{0.074} & \res{0.7673}{0.073}    \\
&		& 6h 39m		& 50h	& 2h 44m		& 11h 10m	& 1h 4m	& 1h 4m\\\hline
NCI1    & 4110 & \res{0.8839}{0.043} & - & \res{0.8851}{0.041} & \res{0.8818}{0.043} & \res{0.8772}{0.047} & \res{0.8811}{0.044}    \\
&	&	39h	&-	& 7h 24m		& 98h	& 2h 1m	& 2h 1m\\\hline
NCI109 & 4127 & \res{0.8728}{0.025} & - & \res{0.8713}{0.025} & \res{0.8670}{0.024}  & \res{0.8713}{0.025} & \res{0.8680}{0.022}  \\
&	&	44h	&-	& 7h 36m		& 99h	& 1h 47m	& 1h 47m\\\hline
CAS     & 4337 & \res{0.9189}{0.012} & - & \res{0.9209}{0.012} & \res{0.9195}{0.11} & \res{0.9161}{0.012} & \res{0.9185}{0.012}    \\
&	&	38h	&-	& 4h 26m		&176h	& 1h 34m		& 1h 34m\\
\hline
\end{tabular}}
\end{table}
